Watch Out for the Bully! Job Interference Study on Dragonfly Network

### 2016-04-03 draft.tex
rewriting abstract and introduction done; tweak format in section 4

The high-radix, low-diameter dragonfly network will be the choice of many computing facilities for building their next generation supercomputers. Preliminary studies show that random placement with adaptive routing should be the rule of thumb to utilize such networks, since random placement with adaptive routing can uniformly distribute traffic over the network and alleviate congestion. However, in this paper, we made the observation that while random placement coupled with adaptive routing can load-balance the network when multiple jobs are running concurrently, they can not guarantee the best performance for each single job. The performance improvement of communication-intensive jobs comes with the sacrifice of less intensive jobs. We refer this phenomenon as ``bully" between concurrently running jobs. We presented a detailed analysis about this ``bully" behavior between jobs at the network level based on a sophisticated, high-fidelity discrete even-driven simulation tool. The observations and analysis presented in this work illuminate a path to the adoption of more efficient job placement policy and routing algorithm for HPC systems with dragonfly network.

>>>>

The high-radix, low-diameter dragonfly network will be the choice of many computing facilities for building their next generation supercomputers. 
Preliminary studies show that random placement with adaptive routing should be the rule of thumb to utilize such networks for it can uniformly distribute traffic and alleviate congestion. 
However, in this paper, we made the observation that while random placement coupled with adaptive routing can load-balance the network, it can not guarantee the best performance for every single job.
We have observed that the performance improvement of communication-intensive jobs comes with the sacrifice of less intensive jobs. 
We identify this ``bully" behavior with the help of detailed network simulation and real application traces and validate the underneath causes.
We propose a hybrid job placement policy that provide either contiguous or random allocation to each application based on its communication intensity. 
With the help of the simulation, we demonstrate that this hybrid job placement policy can lead to performance improvement for less communication-intensive applications.

%We refer this phenomenon as ``bully" between concurrently running jobs. We presented a detailed analysis about this ``bully" behavior between jobs at the network level based on a sophisticated, high-fidelity discrete even-driven simulation tool. The observations and analysis presented in this work illuminate a path to the adoption of more efficient job placement policy and routing algorithm for HPC systems with dragonfly network.

### 2016-04-03 intro.tex
rewriting abstract and introduction done; tweak format in section 4

Recent studies suggest that random job placement coupled with adaptive routing can alleviate local congestion, eliminate hot-spots and reach load-balance for dragonfly networks~\cite{jain-sc14, bhatele-sc11, brandt2014}. They focus on the performance of network without analyzing the performance of each individual application. In this work....

@@

    \item Random placement coupled with adaptive routing can uniformly distribute traffic over the network, avoid hot-spots and alleviate traffic congestion, optimizing the overall performance of the dragonfly network.

    \item The ``bully" behavior occurs between concurrently running jobs on a dragonfly network when random placement and adaptive routing are in use. Communication-intensive jobs obtain performance improvement at the expense of less intensive jobs suffering performance degradation.

    \item Random placement coupled with adaptive routing can improve the performance of communication-intensive jobs by enabling network resource sharing, though it introduces interference between concurrently running jobs.

    \item Contiguous placement coupled with minimal routing can guarantee consistency performance of less communication-intensive jobs by prohibiting network resource sharing, though it results in local network congestion and unbalanced network utilization.

@@

Based on our observations, the optimal job placement and routing solution for the workload running dragonfly system should be specific to the job communication intensity. \emph{The random placement and adaptive routing are preferable to communication-intensive jobs, whereas contiguous placement and minimal routing are in favor of less intensive jobs.} To the best of our knowledge, using parallel application traces from production systems for the study of job interference on dragonfly networks has not been reported so far. We believe the observations and analysis presented in this paper are valuable to a number of communities including HPC computing centers, system software developers, and system administrators. For example, the computing centers should take resource sharing into consideration when choosing proper configurations for building their dragonfly networks. System software developers could design better scheduling algorithms for jobs with distinct communication behavior. System administrators could make more accurate predication about system availability based on job running status under different placement and routing configurations.

>>>>

Recent studies suggest that random job placement coupled with adaptive routing can alleviate local congestion, eliminate hot-spots and reach load-balance for dragonfly networks~\cite{jain-sc14, bhatele-sc11, brandt2014}. These works explore the possible job placement and routing configurations that could optimize the overall network performance.
In this paper, we study the interference among concurrently running applications on dragonfly networks when different job placement and routing policies are in use. 
Our analysis focus on the performance of each application with interference in presence.

%We have observed that when running concurrently on the dragonfly network, applications interfere with each other and the less communication-intensive applications are ``bullied" by the communication-intensive ones. The ``bully" applications get great performance improvement at the cost of performance degradation to the ``bullied" applications. We choose three representative applications from the DOE Design Forward Project~\cite{designforwardwebpage} to study the root cause of the ``bullying" with simulation on CODES, an high-fidelity, packet-level HPC network simulation toolkit~\cite{codes}.

We choose three representative applications from the DOE Design Forward Project~\cite{designforwardwebpage} to analyze the interference among them. Our study is based on the simulation with CODES, an high-fidelity, packet-level HPC network simulation toolkit~\cite{codes}.
For each application, we examine it with two job placement policies and three routing policies.
We make the following interesting observations through extensive simulation.

@@

    \item Concurrently running applications on a dragonfly network interfere with each other when they share network resources. Communication-intensive applications ``bully" their less intensive peers and obtain performance improvement at the expense of less intensive ones suffering performance degradation.

    \item Random placement can improve the performance of communication-intensive applications by enabling network resource sharing, though it introduces interference causing performance degradation.

    \item Contiguous placement can guarantee consistency performance of less communication-intensive applications by prohibiting network resource sharing, though it results in local network congestion and unbalanced network utilization.

@@

Based on the aforementioned key observations, the optimal job placement policy for the workload running dragonfly system should be specific to the application communication intensity. Thus, we propose a hybrid job placement, which can assign the communication-intensive applications with random allocation and less intensive applications with contiguous allocation. Our experiment results show that hybrid job placement can mitigate the performance degradation of less communication-intensive applications while still guarantee the best performance of communication-intensive applications.
%\emph{The random placement and adaptive routing are preferable to communication-intensive jobs, whereas contiguous placement and minimal routing are in favor of less intensive jobs.}

To the best of our knowledge, using parallel application traces from production systems for the study of job interference on dragonfly networks has not been reported so far. We believe the observations and new placement policy presented in this paper are valuable to a number of communities including HPC computing centers, system software developers, and system administrators. For example, the computing centers should take resource sharing into consideration when choosing proper configurations for building their dragonfly networks. System software developers could design better scheduling algorithms for jobs with distinct communication behavior. System administrators could make more accurate predication about system availability based on job running status under different placement and routing configurations.
